---
title: "Homework 5"
author: "Celine Park"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    theme: yeti
editor: visual
execute: 
  message: false
  warning: false
---

**Introduction**: 
Sarrencia, otherwise known as the North American Pitcher plant, are the largest group of carnivorous plants and Sarrencia is one of the largest pitcher plant genera. According to the National Library of Medicine, the leaf extract of Sarrencia is a "traditional medicine for the treatment of type 2 diabetes", and possesses "antimycobacterial activity for the treatment of tuberculosis-like symptoms". Pitcher plants could be useful to predict individual biomass from physiological adaptations because when pitcher plants have evolved to survive in nutrient-poor environments. Pitcher plants "attracts, traps, and digests prey, to absorb their soluble ingredients for growth and reproduction." Predicting biomass for this physiological characteristic could be important because one could study how, depending on the plant's biomass, each pitcher plant could absorb nutrients faster. The question we are thus addressing is how Sarracenia characteristics predict biomass. The hypothesis that we are testing is that ______. 

Loading the libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggeffects)
library(performance)
library(naniar) # or equivalent
library(flextable) # or equivalent
library(car)
library(broom)
library(corrplot)
library(AICcmodavg)
library(GGally)
```

Read in data
```{r}
plant <- read_csv(here("data", "knb-lter-hfr.109.18", "hf109-01-sarracenia.csv")) %>% 
  # make the column names cleaner
  clean_names() %>% 
  # selecting the columns of interest
  select(totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls)
```

**Methods**:
Sampling Methods: 
a. The data that was collected was collected by conducting a manipulative feeding experiment to test morphological and physiological allometric relationships of carnivorous plants when nutrients are not limiting are more similar to allometric relationships of non-carnivorous plants. The effects of prey availibility oh photosynthetic rate, chlorophyll fluorescence, growth, architecture, and foliar nutrient and chlorophyll content were tested.

b. The data was organized with 32 variables and 120 observations. The data was processed through through quarto document, where the relevant columns were selected (totmass, species, feedlevel, sla, chlorophyll, amass, num_lvs, num_phylls). The data was also cleaned by dropping the NA values. 

c. Visualizing missing data:
```{r}
gg_miss_var(plant) + 
  labs(caption = "This is a visualization of the missing data. This visualization shows that there is missing data for chlorophyll, amass, sla, num_phylls, and num_lvs.") +
  theme(plot.caption = element_text(size = 7, hjust=0.5))

```
This shows that the data may be affected due to the missing variables that are in this dataset. Due to the large amount of missing data from chlorophyll, amass, and sla, the precision of the findings may be impacted. 

d. Visualization of Pearson's correlation 

```{r}
plant_subset <- plant %>% 
  drop_na(sla, chlorophyll, amass, num_lvs, num_phylls) #subsetting the data by dropping NAs

# calculate Pearson's r for numerical values only
plant_cor <- plant_subset %>% 
  select(feedlevel:num_phylls) %>% 
  cor(method = "pearson")
  
# creating a correlation plot
corrplot(plant_cor,
         # change the shape of what's in the cells
         method = "ellipse",
         addCoef.col = "black") 
  title(sub="To determine the relationships between numerical variables in our dataset, we calculated Pearsons r and visually represented correlation using a correlation plot. For this plot, positive correlations are displayed in blue, and negative correlations are in red. The color intensity and the size of the circle are proportional to the correlation coefficients.")


  

```
Thus, while analyzing this plot, we can see that the strongest correlation is between amas (mass-based light-saturated photosynthetic rate of youngest leaf) and sla (specific leaf area of youngest fully-expanded leaf produced by the plant). The weakest (negative) correlation is with amass and num_lvs (total number of pitchers and phyllodes produced by each plant). This shows that there is a strong correlation between between the photosynthetic rate of the youngest leaf, with the specific leaf area.  

e. Pairs plot
```{r}
plant_subset %>% 
  select(species:num_phylls) %>% 
  ggpairs(title = "This graph creates a plot of each variable compared against the others. Scatterplots of each pair of numeric variables are drawn on the left, while pearson correlation is displayed on the right. Variable distribution is on the diagonal.") 
  
```


This, while analyzing this plot it shows that while analyzing the density plots along the diagonal, the data is skewed to the right, which indicates a positive skew. The scatter plots show that most of the data is clustered to the left, while the histograms and boxplots show that there are outliers and a variation within the distribution of the variables. All the correlation values are less than |0.5|, which indicates that there is not a strong correlation between the relationships between the variables.


f. Null and full model
To 
```{r}
null <- lm(totmass ~ 1, data = plant_subset)
full <- lm(totmass ~ species + feedlevel + sla + chlorophyll + amass + num_lvs + num_phylls, data = plant_subset)


```
By analyzing the null and full model, we see that  

g. Visual and statistical assumptions checks
```{r}
par(mfrow = c(2, 2))
plot(full)
```







